---
title: 'STA302 Project part 1'
author: "Hilary"
date: "2024-10-11"
output:
  pdf_document: default
  html_document: default
---

```{r warning=FALSE, echo=FALSE, message=FALSE}
# Load the dataset
data <- read.csv("kijiji_rental_ads_4106.csv")

data <- data[, c("Price...", "Date.Posted", "Bedrooms", 
                 "Bathrooms", "Building.Type", "Size..sqft.", "Pet.Friendly")]

colnames(data) <- c("Price", "Days_on_Market", "Number_of_Bedrooms", 
                    "Number_of_Bathrooms", "Building_Type", "Rental_Size", 
                    "Pet_Friendliness")

#insert outlier removal here
# Function to remove outliers using IQR for a given vector

# Function to remove outliers using IQR for given vector, return a logical index
remove_outliers <- function(x) {
  Q1 <- quantile(x, 0.25, na.rm = TRUE)
  Q3 <- quantile(x, 0.75, na.rm = TRUE)
  IQR <- Q3 - Q1
  lower_bound <- Q1 - 1.5 * IQR
  upper_bound <- Q3 + 1.5 * IQR
  return(x >= lower_bound & x <= upper_bound)
}

# Function to remove outliers using IQR for a given numeric vector and return a logical index
remove_outliers <- function(x) {
  if (is.numeric(x)) {  # Check if the column is numeric
    Q1 <- quantile(x, 0.25, na.rm = TRUE)
    Q3 <- quantile(x, 0.75, na.rm = TRUE)
    IQR <- Q3 - Q1
    lower_bound <- Q1 - 1.5 * IQR
    upper_bound <- Q3 + 1.5 * IQR
    return(x >= lower_bound & x <= upper_bound)
  } else {
    return(rep(TRUE, length(x)))  # If it's not numeric, then return TRUE for all entries
  }
}

index <- remove_outliers(data$Price) &
         remove_outliers(data$Days_on_Market) &
         remove_outliers(data$Number_of_Bedrooms) &
         remove_outliers(data$Number_of_Bathrooms) &
         remove_outliers(data$Rental_Size)

windex <- remove_outliers(data$Days_on_Market)

data_clean <- data[index, ]

# Remove commas and convert Rental_Size to numeric
data_clean$Rental_Size <- gsub(",", "", data_clean$Rental_Size) 
data_clean$Rental_Size <- as.numeric(data_clean$Rental_Size)    

# Clean Number_of_Bedrooms by replacing text and converting to numeric
data_clean$Number_of_Bedrooms <- gsub("Bachelor/Studio", "0", data_clean$Number_of_Bedrooms)  
data_clean$Number_of_Bedrooms <- as.numeric(data_clean$Number_of_Bedrooms)  

# Convert Number_of_Bathrooms to numeric
# Replace "Not Available" or other texts with NA to handle non-numeric entries
data_clean$Number_of_Bathrooms <- gsub("Not Available", NA, data_clean$Number_of_Bathrooms)  
data_clean$Number_of_Bathrooms <- as.numeric(data_clean$Number_of_Bathrooms) 

#convert date format to days
fixed_date <- ymd("2024-10-10")
data_clean$Days_on_Market <- as.Date((data_clean$Days_on_Market))
data_clean$Days_on_Market <- as.numeric(difftime(fixed_date, data_clean$Days_on_Market, units = "days"))

data_clean <- data_clean |> filter(!is.na(Days_on_Market), !is.na(Number_of_Bedrooms), !is.na(Number_of_Bathrooms), !is.na(Rental_Size))
```


```{r}

```



```{r warning=FALSE, echo=FALSE, message=FALSE}

library(ggplot2)
library(patchwork)

# Fit the linear model
model1 <- lm(Price ~ Days_on_Market + Number_of_Bedrooms + Number_of_Bathrooms + Rental_Size, data = data_clean)

# Create ggplot residual plots
# Residuals vs Fitted Values
plot_residuals_vs_fitted <- ggplot(data.frame(Fitted = fitted(model1), Residuals = residuals(model1)), 
                                   aes(x = Fitted, y = Residuals)) +
  geom_point(color = "blue", size = 0.5) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(title = "Residuals vs Fitted Values", x = "Fitted Values", y = "Residuals") +
  theme_minimal()

# Residuals vs Days on Market
plot_residuals_vs_days <- ggplot(data_clean, aes(x = Days_on_Market, y = residuals(model1))) +
  geom_point(color = "blue", size = 0.5) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(title = "Residuals vs Days on Market", x = "Days on Market", y = "Residuals") +
  theme_minimal()

# Residuals vs Number of Bedrooms
plot_residuals_vs_bedrooms <- ggplot(data_clean, aes(x = Number_of_Bedrooms, y = residuals(model1))) +
  geom_point(color = "blue", size = 0.5) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(title = "Residuals vs Number of Bedrooms", x = "Number of Bedrooms", y = "Residuals") +
  theme_minimal()

# Residuals vs Number of Bathrooms
plot_residuals_vs_bathrooms <- ggplot(data_clean, aes(x = Number_of_Bathrooms, y = residuals(model1))) +
  geom_point(color = "blue", size = 0.5) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(title = "Residuals vs Number of Bathrooms", x = "Number of Bathrooms", y = "Residuals") +
  theme_minimal()

# Residuals vs Rental Size
plot_residuals_vs_rental <- ggplot(data_clean, aes(x = Rental_Size, y = residuals(model1))) +
  geom_point(color = "blue", size = 0.5) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(title = "Residuals vs Rental Size", x = "Rental Size", y = "Residuals") +
  theme_minimal()

# Normal Q-Q Plot
plot_qq <- ggplot(data.frame(Residuals = residuals(model1)), aes(sample = Residuals)) +
  stat_qq(color = "black", shape=1) +
  stat_qq_line(color = "red") +
  labs(title = "Normal Q-Q Plot", x = "Theoretical Quantiles", y = "Sample Quantiles") +
  theme_minimal()

# Arrange the plots using patchwork
combined_plot <- (
  (plot_residuals_vs_days | plot_residuals_vs_bedrooms) /
  (plot_residuals_vs_rental | plot_residuals_vs_bathrooms) /
  (plot_residuals_vs_fitted | plot_qq) 
)

# Display the combined plot
print(combined_plot)

```

```{r warning=FALSE, echo=FALSE, message=FALSE}
# Fit the linear model
model2 <- lm(Price ~ Days_on_Market + Number_of_Bedrooms + Building_Type + Rental_Size, data = data_clean)

# Create ggplot residual plots
# Residuals vs Fitted Values
plot_residuals_vs_fitted <- ggplot(data.frame(Fitted = fitted(model2), Residuals = residuals(model2)), 
                                   aes(x = Fitted, y = Residuals)) +
  geom_point(color = "blue", size = 0.5) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(title = "Residuals vs Fitted Values", x = "Fitted Values", y = "Residuals") +
  theme_minimal()

# Residuals vs Days on Market
plot_residuals_vs_days <- ggplot(data_clean, aes(x = Days_on_Market, y = residuals(model1))) +
  geom_point(color = "blue", size = 0.5) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(title = "Residuals vs Days on Market", x = "Days on Market", y = "Residuals") +
  theme_minimal()

# Residuals vs Number of Bedrooms
plot_residuals_vs_bedrooms <- ggplot(data_clean, aes(x = Number_of_Bedrooms, y = residuals(model1))) +
  geom_point(color = "blue", size = 0.5) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(title = "Residuals vs Number of Bedrooms", x = "Number of Bedrooms", y = "Residuals") +
  theme_minimal()

# Residuals vs Building Type
plot_residuals_vs_type <- ggplot(data_clean, aes(x = Building_Type, y = residuals(model2))) +
  geom_boxplot(aes(color = Building_Type)) + 
   
  labs(title = "Residuals vs Building Type", x = "Building Type", y = "Residuals") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 8),
    axis.text.y = element_text(size = 10),
    axis.title = element_text(size = 12),
    ,legend.position = "none")

# Residuals vs Rental Size
plot_residuals_vs_rental <- ggplot(data_clean, aes(x = Rental_Size, y = residuals(model2))) +
  geom_point(color = "blue", size = 0.5) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(title = "Residuals vs Rental Size", x = "Rental Size", y = "Residuals") +
  theme_minimal()

# Normal Q-Q Plot
plot_qq <- ggplot(data.frame(Residuals = residuals(model2)), aes(sample = Residuals)) +
  stat_qq(color = "black", shape=1) +
  stat_qq_line(color = "red") +
  labs(title = "Normal Q-Q Plot", x = "Theoretical Quantiles", y = "Sample Quantiles") +
  theme_minimal()

# Arrange the plots using patchwork
combined_plot2 <- (
  (plot_residuals_vs_days | plot_residuals_vs_bedrooms) /
  (plot_residuals_vs_rental ) /
  (plot_residuals_vs_fitted | plot_qq)
)

building_plot <- ((plot_residuals_vs_type))

# Display the combined plot
print(combined_plot2)
print(building_plot)
```

1. Residuals vs. Fitted Values
In the plot, residuals  are plotted against the fitted values . The residuals do not show a clear pattern but cluster slightly, suggesting the linearity assumption is fitted. 
2. Histogram of Residuals
The histogram shows the distribution of residuals. While the residuals appear close to normal, small deviations affect confidence intervals and p-values. If deviations from normality are significant, we can apply a Box-Cox to help with analysis.
3. Normal Q-Q Plot
The residuals mostly follow the reference line, but there are slight deviations at the tails. This suggests that the residuals are approximately normal, but the deviations at the tails might indicate some outliers.

```{r warning=FALSE, echo=FALSE, message=FALSE}
library(MASS)
# Visualize the Box-Cox transformation
boxcox(model, lambda = seq(-2, 2, by = 0.1))  # Adjust lambda range if needed
```
